---
title: "STAT 757 Assignment 8"
subtitle: "DUE 4/22/2018 11:59PM"
author: "AG Schissler"
date: "2/14/2018"
output: pdf_document
fig_caption: true
exclude: true
caption: true
references:
- id: sheather2009
  title: A modern approach to regression with R
  author:
  - family: Sheather
    given: Simon
  publisher: Springer Science \& Business Media
  type: book
  issued:
    year: 2009
---

## Instructions [20 points]

Modify this file to provide responses to the Ch.8 Exercises in @sheather2009. You can find some helpful code here: [http://www.stat.
tamu.edu/~sheather/book/docs/rcode/Chapter8.R](http://www.stat.tamu.edu/~sheather/book/docs/rcode/Chapter8.R). Also address the project milestones indicated below. Please email **both** your .Rmd (or roxygen .R) and one of the following either .HTML, .PDF, or .DOCX using the format SURNAME-FIRSTNAME-Assignment8.Rmd and SURNAME-FIRSTNAME-Assignment8.pdf.

## Exercise 8.3.4 [60 points] 

### Part A

The marginal model plots show nice agreement for $x2, x5, \hat{y}$. However, $x1,x4$ show poor agreement, especially in the tails. Therefore, the model is invalid for these data as specified.

### Part B

Both $x1,x4$ are right-skewed. Following the guidelines by Sheather, I would include $log(x1),log(x4)$ in the model to account for this.

### Part C

All the marginal model plots show nice agreement; I conclude the model is valid.

### Part D

When a patient has a family history of heart disease ($x3=1$), the log-odds of heart disease is increased by $\hat{\beta}_3=0.941$. This corresponds to an odds ratio of $exp(\hat{\beta}_3)= 2.5625$.

### Appendix

The appendix contains the diagnostics for the model posed in Part B above.

```{r}
myDir <- "~/OneDrive - University of Nevada, Reno/Teaching/STAT_757/Sheather_data/Data/"
dat <- read.delim(file.path(myDir,"HeartDisease.csv"), sep = ",")
str(dat)
```

```{r}
m2 <- glm(HeartDisease ~ x1 + log(x1) + x2 + x3 + x4 + log(x4) + x5,
          family=binomial() , data=dat )
summary(m2)
```

```{r, fig.height=7}
library(alr3)
par(mfrow=c(3,3))
mmp(m2,dat$x1)
mmp(m2,log(dat$x1))
mmp(m2,dat$x2)
mmp(m2,dat$x4)
mmp(m2,log(dat$x2))
mmp(m2,dat$x5)
mmp(m2,m2$fitted.values,xlab="Fitted Values")
```

## Project milestones [20 points]

1. Finalize your analysis.
2. Draft your methods section of the written report.

## References
