---
title: "STAT 757 Assignment 5 Solutions"
subtitle: "DUE 4/01/2018 11:59PM"
author: "AG Schissler"
date: "2/14/2018"
output: pdf_document
fig_caption: true
exclude: true
caption: true
references:
- id: sheather2009
  title: A modern approach to regression with R
  author:
  - family: Sheather
    given: Simon
  publisher: Springer Science \& Business Media
  type: book
  issued:
    year: 2009
---

## Instructions [20 points]

Modify this file to provide responses to the Ch.5 Exercises in @sheather2009. You can find some helpful code here: [http://www.stat.
tamu.edu/~sheather/book/docs/rcode/Chapter5.R](http://www.stat.tamu.edu/~sheather/book/docs/rcode/Chapter5.R). Also address the project milestones indicated below. Please email **both** your .Rmd (or roxygen .R) and one of the following either .HTML, .PDF, or .DOCX using the format SURNAME-FIRSTNAME-Assignment5.Rmd and SURNAME-FIRSTNAME-Assignment5.pdf.

```{r, setPath, echo = T}
data_dir <- "/Users/alfred/OneDrive - University of Nevada, Reno/Teaching/STAT_757/Sheather_data/Data"

```

## Exercise 5.4.2

```{r, HoustonChronicle, echo = T}

reading <- read.csv(file.path(data_dir, "HoustonChronicle.csv"),header=TRUE)
str(reading)
head(reading)
## clean up names
names(reading) <- gsub("X\\.", "", names(reading))
## "factorize" year
reading$Year <- factor(reading$Year)

```

```{r, reading_plot, echo = T, fig.height = 5}
require(ggplot2)
p1 <- ggplot(data = reading, aes(x = reading$Low.income.students, y = Repeating.1st.Grade, shape = Year, color = Year))
p1 + geom_point()
p1 + geom_point() + stat_smooth(method = "lm", col = "blue")
```
The Year category seems fairly randomly disributed across data points. But this may be something that is hard to see by eye.

### 5.4.2 Part a

This is a the analysis of covariance (ANCOVA) scenario with co-incident lines.

```{r, readingA, echo = T, fig.height = 5}
fit1 <- lm(Repeating.1st.Grade ~ Low.income.students, data = reading)
summary(fit1)
par(mfrow=c(2,2))
plot(fit1)
```
Looks like a decent fit and the slope is significantly greater than 0. 

### 5.4.2 Part b

This scenario ignores the low-income percentage.

```{r, readingB, echo = T, fig.height = 5}
fit2 <- lm(Repeating.1st.Grade ~ Year, data = reading)
summary(fit2)
par(mfrow=c(2,2))
plot(fit2)
```
This amounts to a t-test on the different means between the Years. There is not much evidence of an increase when not accounting for low-income status.

### 5.4.2 Part c

The scenario below assumes an additive effect of Year (parallel lines).

```{r, readingB2, echo = T, fig.height = 5}
fit3 <- lm(Repeating.1st.Grade ~ Year + Low.income.students, data = reading)
summary(fit3)
par(mfrow=c(2,2))
plot(fit3)
```

This scenario is the most general ANCOVA.

```{r, readingC, echo = T, fig.height = 5}
full <- lm(Repeating.1st.Grade ~ Low.income.students + Year + Low.income.students:Year, data = reading)
summary(full)
par(mfrow=c(2,2))
plot(full)
```

There is little evidence that the association between the percentage low-income students and percentage reading failure rates differs between the years 1994 and 2004. Let's do an ANOVA to compare models:

```{r, anova_5_4_2}
## check against only parallel lines model
anova(fit3, full)
## check against only Year model
anova(fit2, full) 
## check against only low income model
anova(fit1, full) 

```
Based on this it is reasonable to not consider the Year in the analysis. It does however show including the an interaction with low-income is better than leaving it out entirely (Year-only model). My final analysis would be that the the association does not change with year, but there is a "positive" association between the percentage of low-income students and higher reading test failure rates. The final model is plotting below:

```{r, reading_plot_final, echo = T, fig.height = 5}
require(ggplot2)
p1 <- ggplot(data = reading, aes(x = reading$Low.income.students, y = Repeating.1st.Grade))
p1 + geom_point(aes(shape = Year, color = Year)) + stat_smooth(method = "lm", col = "blue")
```

## Exercise 5.4.3

```{r, latour, echo = T, fig.height = 5}

latour <- read.table(file.path(data_dir, "Latour.txt"), header=TRUE)
head(latour)
str(latour)

#Figure 5.8 on page 149
y <- latour$Rain
par(mfrow=c(1,1))
plot(latour$EndofHarvest, latour$Quality,pch=y+1,col=y+1,xlab="End of Harvest (in days since August 31)")
abline(lsfit(latour$EndofHarvest[y==0],latour$Quality[y==0]),lty=1,col=1)
abline(lsfit(latour$EndofHarvest[y==1],latour$Quality[y==1]),lty=2,col=2)
legend(23, 2.5,legend=c("No","Yes"),pch=1:2,col=1:2,title="Rain at Harvest?")

```

### Exercise 5.4.3 Part A

The plot suggests differing slopes and intercepts when considering the rate of change in wine quality and days after August 31 for years with rain at harvest. Note that the y-intercept is not illustrated in the plot. As instructed in the prompt, first let's fit the most general model:

```{r, latourA, echo = T}
#Regression output on page 148
mfull <- lm(Quality ~ EndofHarvest + Rain + Rain:EndofHarvest, data = latour)
summary(mfull)

#Regression output on page 149
mreduced <- lm(Quality ~ EndofHarvest + Rain, data = latour)
summary(mreduced)
anova(mreduced,mfull)

```
By inspecting the anova analysis, it is clear that the impact of a rain on the rate of change between the wine quality and later harvest is statistically significant.

### Exercise 5.4.3 Part B

To estimate the days delay for one point decrease in wine quality, we follow the description on the top of page 141 in @sheather2009 and note that:

\begin{align*}
E(Y | d=0) = \beta_0 + \beta_1 x & \Rightarrow \\ 
E(\Delta Y | d=0) &= \beta_0 + \beta_1 \Delta x\\
-1 &= \beta_0 + \beta_1 \Delta x \\
\frac{-1-\beta_0}{\beta_1} &= \Delta x
\end{align*}

In the absence of rain, we estimate the days delayed to correspond with a point decrease in quality to be
```{r, latourB1, echo = T}
unname((-1-coefficients(mfull)[1])/coefficients(mfull)[2])
```

Similarly,

\begin{align*}
E(Y | d=1) = \beta_0 + \beta_2 + (\beta_1 + \beta_3) x & \Rightarrow \\ 
E(\Delta Y | d=1) &= \beta_0 + \beta_2 + (\beta_1 + \beta_3) \Delta x \\
-1 &= \beta_0 + \beta_2 + (\beta_1 + \beta_3) \Delta x \\
\frac{-1-\beta_0 - \beta_2}{\beta_1 + \beta_3} &= \Delta x
\end{align*}


In the presence of rain, we estimate the days delayed to correspond with a point decrease in quality to be
```{r, latourB2, echo = T}
unname((-1-coefficients(mfull)[1] - coefficients(mfull)[3])/(coefficients(mfull)[2] + coefficients(mfull)[4]))
```

## Project milestones [20 points]

1. Perform an exploratory data analysis:
 * Numerically summarize the variables.
 * Make plots and explore relationships between variables.
 * Identify any strange points or anything else that doesn't make sense.
2. Begin to think about how to model the relationships in your data.

## References
