---
title: "STAT 757 Assignment 2 Solutions"
subtitle: "DUE 2/18/2018 11:59PM"
author: "AG Schissler"
date: "2/1/2018"
output: pdf_document
fig_caption: true
exclude: true
references:
- id: sheather2009
  title: A modern approach to regression with R
  author:
  - family: Sheather
    given: Simon
  publisher: Springer Science \& Business Media
  type: book
  issued:
    year: 2009
---

# 2.8.1 [10 points]

```{r, explore_playbill, fig.cap = "Note the striking linear relationship between box office results lagged by one week."}
## change the line below to your path!
my_data_path <- "~/OneDrive - University of Nevada, Reno/Teaching/STAT_757/Sheather_data/Data"
playbill <- read.csv(file.path(my_data_path,"playbill.csv"),header=TRUE) 

# Figure 2.6 on page 38
plot(playbill$LastWeek,playbill$CurrentWeek,xlab="Gross box office results previous week",
ylab="Gross box office results current week")
```

## 2.8.1 Part A

```{r, parta}
lm_fit <- lm(formula = CurrentWeek ~ LastWeek, data = playbill)

## uncomment below to see the available functions for the lm fit object
## attributes(lm_fit)

lm_summary <- summary(lm_fit)
## attributes(lm_summary)

## 95% confidence interval using the output of lm
t_star <- qt(0.975, df = nrow(playbill) - 2)
B1_hat <- coefficients(lm_summary)["LastWeek", "Estimate"]
B1_se <- coefficients(lm_summary)["LastWeek", "Std. Error"]

margin_of_error <- t_star * B1_se
## technically half the interval's length (not the normal definition of margin of error)

print("A 95% confidence interval")
c(lower = B1_hat - margin_of_error, upper = B1_hat + margin_of_error)

## the easy way:
confint(lm_fit)
```
Since the interval contains the value 1, it's plausible that $\beta_{1} = 1$.

## 2.8.1 Part B
```{r, partb}
B0_hat <- coefficients(lm_summary)["(Intercept)", "Estimate"]
B0_se <- coefficients(lm_summary)["(Intercept)", "Std. Error"]
B0_null = 1e4

## build test statistic, see section 2.2.3
t_partb <- ( B0_hat - B0_null ) / B0_se
t_partb

## two-sided test
2*pt(q = t_partb, df = nrow(playbill) - 2)
```
There is little evidence to conclude that $\beta_{0}$ is different from 10000.

## 2.8.1 Part C
```{r, partc}
## y_hat vector is found by the predict function
(y_hat <- predict(object = lm_fit))

## now predict a new value
x_star <- 400000
unname(coefficients(lm_fit)[1] + coefficients(lm_fit)[2]*x_star)

## or use the predict function
x_star_data <- data.frame(LastWeek = 400000) ## change data structure make function work
(y_hat_star <- predict(object = lm_fit, newdata = x_star_data))

## compute a 95% prediction interval around the prediction for 40000
n <- nrow(playbill)
t_star <- qt(0.975, df = n - 2)

## make a prediction using Equations (2.15, 2.16, 2.17)
x_mean <- mean(playbill$LastWeek)
SXX <- sum( (playbill$LastWeek - x_mean)^2 )

## estimate sigma
RSS <- sum((playbill$CurrentWeek - y_hat)^2)
S <- sqrt( RSS / (n-2))
y_hat_star_se <- S * sqrt(1 + n^(-1) + ( (x_star - x_mean)^2 / SXX ) )

margin_of_error <- t_star * y_hat_star_se

print("A 95% prediction interval when x = 400000")
c(lower = y_hat_star - margin_of_error, upper = y_hat_star + margin_of_error)

## or use the predict function
predict(object = lm_fit, x_star_data, interval = "prediction")

## now predict a new value
x_star <- 400000
unname(coefficients(lm_fit)[1] + coefficients(lm_fit)[2]*x_star)

## or use the predict function
x_star_data <- data.frame(LastWeek = 400000) ## change data structure make function work
(y_hat_star <- predict(object = lm_fit, newdata = x_star_data))

## compute a 95% prediction interval around the prediction for 40000
n <- nrow(playbill)
t_star <- qt(0.975, df = n - 2)

## make a prediction using Equations (2.15, 2.16, 2.17)
x_mean <- mean(playbill$LastWeek)
SXX <- sum( (playbill$LastWeek - x_mean)^2 )

## estimate sigma
RSS <- sum((playbill$CurrentWeek - y_hat)^2)
S <- sqrt( RSS / (n-2))
y_hat_star_se <- S * sqrt(1 + n^(-1) + ( (x_star - x_mean)^2 / SXX ) )

margin_of_error <- t_star * y_hat_star_se

print("A 95% prediction interval when x = 400000")
c(lower = y_hat_star - margin_of_error, upper = y_hat_star + margin_of_error)

## or use the predict function
predict(object = lm_fit, x_star_data, interval = "prediction")
```

No, $450,000 is well outside the prediction interval.

## 2.8.1 Part D
Predicting the current week sales from the previous week is a somewhat reasonable strategy. Reformulating the question into terms of the model, predicting current sales using the exact previous sales is assuming that $\beta_{0} = 0$ and $\beta_{1} = 1$. From our investigation, we believe that $\beta_{0}$ could be 0 and $\beta_{1}$ could be one. However, if one continually predicts the current week from the last week then the obvious and common sense trend of decreasing sales is not being recognized. Indeed our $\hat{\beta_{1}} < 1$, hinting at that logic. More data or a different model could produce a better prediction rule. 

# Exercise 2.8.2 [10 points]

```{r, explore_indicators, fig.cap = "The relationship between the variables is noisy and perhaps only weakly associated."}
## change the line below to your path!
my_data_path <- "~/OneDrive - University of Nevada, Reno/Teaching/STAT_757/Sheather_data/Data"
indicators_dat <- read.table(file.path(my_data_path,"indicators_v2.txt"), header=T, sep = "\t")

# Figure 2.6 on page 38
plot(indicators_dat$LoanPaymentsOverdue,indicators_dat$PriceChange, xlab="Percentage of mortgage loans 30 days or more overdue in latest quarter",
ylab="Percentage change in average price from Jul 2006 to Jul 2007.")
```

## 2.8.2 Part A

```{r, 2parta}
lm_fit <- lm(formula = PriceChange ~ LoanPaymentsOverdue, data = indicators_dat)
## uncomment below to see the available functions for the lm fit object
## attributes(lm_fit)
lm_summary <- summary(lm_fit)
## attributes(lm_summary)
## the easy way:
confint(lm_fit)
```

Since the 95% confidence interval lies entirely the negative side of 0, there is statistically significant evidence that $\beta_{1} \ngeq 0$.

## 2.8.2 Part B

```{r, 2partb}
x_new <- data.frame(LoanPaymentsOverdue = 4)
predict(object = lm_fit, x_new, interval = "confidence")
```

Note that the question is referring to the parameter $E(Y | X=4$), not the random variable $\hat{y}$. Regression parameters estimates are less variable than the prediction estimates associated with the random variable. See `?predict.lm` for more reading. This more narrow interval does not contain 0 and so the average value of $Y | X = 4$ is likely to be less than 0.

# Exercise 2.8.3 [10 points]

```{r, explore_invoices, fig.cap = "Fairly linear relationship between x and y."}
## change the line below to your path!
my_data_path <- "~/OneDrive - University of Nevada, Reno/Teaching/STAT_757/Sheather_data/Data"
invoices_dat <- read.table(file.path(my_data_path,"invoices.txt"),header=TRUE) 

# Figure 2.6 on page 38
plot(invoices_dat$Invoices,invoices_dat$Time ,xlab="Number of Invoices for the day",
ylab="Processing time in hours")
```

## 2.8.3 Part A

```{r, 3parta}
lm_fit <- lm(formula = Time ~ Invoices, data = invoices_dat)
## uncomment below to see the available functions for the lm fit object
## attributes(lm_fit)
(lm_summary <- summary(lm_fit))
## attributes(lm_summary)
## the easy way:
confint(lm_fit)[1,]
```

## 2.8.3 Part B

```{r, 3partb}
B1_hat <- coefficients(lm_summary)["Invoices", "Estimate"]
B1_se <- coefficients(lm_summary)["Invoices", "Std. Error"]
B1_null = 0.01

## build test statistic, see section 2.2.3
t_partb <- ( B1_hat - B1_null ) / B1_se
t_partb

## two-sided test (careful with the direction here)
2*pt(q = t_partb, df = nrow(invoices_dat) - 2, lower.tail = F)
```
There is some evidence to believe that $\beta_{1}$ is different from 0.01 hours/invoice, but not at conventional significance levels.

## 2.8.3 Part C

```{r, 3partc}
x_star_dat <- data.frame(Invoices = 130)
predict(lm_fit, newdata = x_star_dat, interval = "prediction", level = 0.95)
```
The predicted value of the number of hours is 2.1 hours with a 95% prediction interval between 1.4 and 2.8 hours.

# Exercise 2.8.4 [10 points]

## 2.8.4 Part A

We proceed by minimizing the residual sum of squares ($RSS$). 
\begin{align*}
0 \equiv \frac{\partial RSS}{\partial b} &= \frac{\partial}{\partial b} \sum_{i=1}^n (y_i - bx_i)^2 \\
=& -2 \sum_{i=1}^n (y_i - bx_i)*x_i \\
=& -2 [\sum_{i=1}^n x_iy_i - b\sum_{i=1}^n x_i^2] \Rightarrow \\
& b \sum_{i=1}^n x_i^2 = \sum_{i=1}^n x_iy_i 
\end{align*}

Which gives

\begin{equation}
\hat{\beta} = \frac{\sum_{i=1}^n x_iy_i}{\sum_{i=1}^n x_i^2}
\end{equation}

Since the function is quadratic in $b$ this is a global minimum.

## 2.8.4 Part B

First note that $\hat{\beta} = \sum_{i=1}^{n} c_{i} y_{i}$ where $c_{i} = \frac{x_{i}}{\sum_{i=1}^{n} x_{i}^{2}}$. Conditional on the observed $X$ this $c_i$ is a constant.

### (i)

Then in a similar fashion to @sheather2009 Section 2.7.1, we proceed as follows:
\begin{align*}
E(\hat{\beta} | X) &= E[\sum_{i=1}^{n} c_{i} y_i| X = x_i] \\
&= \sum_{i=1}^{n} c_{i} E[ y_i| X = x_i] \\
&= \sum_{i=1}^{n} c_{i} E[ \beta x_i + \epsilon_i] \\
&= \sum_{i=1}^{n} c_{i}[ E(\beta x_i) + E(\epsilon_i)]\\
&= \sum_{i=1}^{n} c_{i}[ \beta x_i + 0] \\
&= \beta \sum_{i=1}^{n} c_{i}x_i \\
&= \beta \sum_{i=1}^{n} \frac{x_{i}x_i}{\sum_{i=1}^{n} x_{i}^{2}} \\
&= \beta
\end{align*}

### (ii)

\begin{align*}
Var(\hat{\beta} | X) &= Var[\sum_{i=1}^{n} c_{i} y_i| X = x_i] \\
&= \sum_{i=1}^{n} c_{i}^2 Var[ y_i| X = x_i] \\
&= \sum_{i=1}^{n} c_{i}^2 \sigma^2 \\
&= \sigma^2 \sum_{i=1}^{n} c_{i}^2  \\
&= \sigma^2 \sum_{i=1}^{n} \left(\frac{x_{i}}{\sum_{i=1}^{n} x_{i}^{2}}\right)^2  \\
&= \sigma^2 \sum_{i=1}^{n} \frac{x_{i}^2}{(\sum_{i=1}^{n} x_{i}^{2})^2}  \\
&= \sigma^2 \frac{\sum_{i=1}^{n} x_{i}^2}{(\sum_{i=1}^{n} x_{i}^{2})^2}  \\
&= \frac{\sigma^2}{\sum_{i=1}^{n} x_{i}^{2}}
\end{align*}

### (iii)
Using parts (i) and (ii) above and since $\hat{\beta} | X$ is a linear combination of normal randoms variables by construction (namely, the $Y_{i} | X_{i}$), $\hat{\beta} | X$ must be normally distributed with the indicated mean and standard deviation.

# Exercise 2.8.5 [10 points]
Statement (d) is true. Total variation in $Y$, denoted $SST$ or $SYY$, can be decomposed into residual sum of squares $RSS$ and regression sum of squares $SS_{reg}$ and each component of the variation competes with one another. So it would make sense that the directionality of the statements must contradict. To see the direction, the sum of squared residuals in Model 1 are small (differences between the regression line and the points) while deviations from $\bar{y}$ (imagine a horizontal line around y = 5) that contribute to $SS_{reg}$ are large. Exactly the opposite is true for Model 2. 

# Exercise 2.8.6 [10 points]

## Part a

\begin{align*}
(y_i-\hat{y}_i) &= y_i - \hat{\beta}{_0} - \hat{\beta}{_1}x_i \\
&= y_i - (\bar{y} - \hat{\beta}{_1}\bar{x}) - \hat{\beta}{_1}x_i \\
&= (y_i - \bar{y}) - \hat{\beta}{_1}(x_i - \bar{x}) 
\end{align*}

## Part b

\begin{align*}
(\hat{y}_i - \bar{y}) &=  (\hat{\beta}{_0} + \hat{\beta}{_1}x_i) - \bar{y} \\
&=  [(\bar{y} - \hat{\beta}{_1}\bar{x}) + \hat{\beta}{_1}x_i] - \bar{y} \\
&=  \hat{\beta}{_1}(x_i - \bar{x})
\end{align*}

## Part c

\begin{align*}
\sum_{i=1}^n (y_i-\hat{y}_i)(\hat{y}_i - \bar{y}) &= \sum_{i=1}^n ([(y_i - \bar{y}) - \hat{\beta}{_1}(x_i - \bar{x})][\hat{\beta}{_1}(x_i - \bar{x})]) \\
&= \hat{\beta}{_1} \sum_{i=1}^n[(y_i - \bar{y})(x_i-\bar{x})] - \hat{\beta}_1^2 \sum_{i=1}^n (x_i-\bar{x})^2 \\
&= \frac{SXY}{SXX} * SXY - \left(\frac{SXY}{SXX}\right)^2 * SXX \\
&= \frac{(SXY)^2 - (SXY)^2 }{SXX} = 0
\end{align*}

This shows that $SST = RSS + SS_{reg}$ by simply expanding $SST$ and seeing the result in part c (the cross term) is zero. 

# Exercise 2.8.7 [10 points]

The confidence interval shown illustrates the variation in inferences on the regression parameters at every observed $x$ value. This is different from variation that can be observed for the random variable $Y$ (condition on $X = x$). The regression paramaters are the expected value for $Y$ and as such the estimation of confidence intervals can be very narrow when there are many data points. Nothing in the theory guarantees any points to be in the mean's CI.

## References
